Evaluate PEER in each cohort and brain region

### 2021-02-18 combined_analysis/eval_resid_in_each.Rmd
add more summary info

# # HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
           dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
           # dplyr::filter(Region %in% c('DLPFC')) %>%
         dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
         # dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)

@@

# drop individuals who where sequenced twice here
tab = table(METADATA$Individual_ID)

dropSamples = sapply( names(tab[tab > 1]), function(id){

  idx = which(METADATA$Individual_ID == id)
  df = METADATA[idx,]

  # remove individuals that have less then max reads
  i = which(df$MappedReads < max(df$MappedReads))
  rownames(METADATA)[idx[i]]
  } )
METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
```

>>>>

# HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

# NIMH-HBCC-5
counts.HBCC5 = fread(synGet('syn21867762')$path, data.table=FALSE)
rownames(counts.HBCC5) = counts.HBCC5$Geneid
counts.HBCC5 = counts.HBCC5[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4, counts.HBCC5)
```

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
  dplyr::filter(ageOfDeath > 17)

@@

# When individuals were sequenced multiple times from the same brain region
#   for technical replicates or QC failure,
# Keep sample with most reads
METADATA = METADATA %>%
  group_by(Individual_ID, Region) %>% 
  slice_max(MappedReads)
```
